# Personal Microbial Cloud

## Taxon Analysis

Read in prepared workspace and load packages.

```{r}
load('../data/pb_data.RData')
library(vegan)
library(labdsv)
library(ape) 
library(xtable)
source('../data/pb_functions_for_manuscript.R')
```



How do samples fall into groups? I.e., do we have balance in treatments?

```{r}
table(pb.map$person, pb.map$sampleType)
table(pb.map$location)
table(pb.map$duration)
table(pb.map$sampleType)
```




Quick refrerence vector of names for figures and whatnot. Some genus, families, etc are empty, so search for the finest taxonomic level and use that name. `cons` is in functions script sourced above. 

```{r}
consensus <- apply(taxo[, 1:6], 1, cons)
consensus[1:10]
```


Generate a list of the 10 most abundant OTUs overall, but also tack on OTUs that are among the 10 most abundant in any particular group of samples. Then a matching version of the mapping table. 

```{r}
tops <- names(rev(sort(colSums(pb.1000))))[1:10]
for (i in 1:length(gr2)) {
  tops.plus <- names(rev(sort(colSums(pb.1000[gr2[[i]], ]))))[1:10]
  tops <- unique(c(tops, tops.plus))
	}
tops <- tops[rev(order(taxo[tops, 7]))]
taxo.tops <- taxo[tops, ]
length(tops)
consensus[tops]
```


Create a reordered map and otu table. 

```{r}
pb.order <- data.frame(pb.1000[unlist(gr2), tops])
pb.order$gr.num <- c(rep(gr.name, c(unlist(lapply(gr2, length)))))
map.order <- pb.map[row.names(pb.order), ]
pb.order$location <- factor(map.order$location, levels=c('occ', 'unocc'))
```


Make new trimmed OTU table and assess with boxplots. Then send the statistically interesting otus to figures and tables script. 

```{r}
pb.order <- pb.order[map.order$duration == '4', ]
# for (i in 1:(ncol(otus.to.barplot.tmp)-2)) {  
# 	boxplot(otus.to.barplot.tmp[, i] ~ factor(otus.to.barplot.tmp$gr.num), las=2, 
# 		main=consensus[gsub('X', '', names(otus.to.barplot.tmp)[i])], sub=i)
# 	readline("Press return for next plot")
# }
pb.order <- pb.order[, c(1, 2, 3, 4, 5, 6, 8, 10, 12, 18, 21, ncol(pb.order)-1, ncol(pb.order))]
names(pb.order) <- gsub('X', '', names(pb.order))
consensus[names(pb.order)]
rm(cons, Evenness, gr.name, gr2, tops.plus, makeTaxo) 
```



Now for indicator analysis - this takes a long time and lots of memory! This uses the `indval()` function in the `labdsv` package. This gives output for all OTUs, the vast majority of which are unimportant to analysis. So there are a few steps to cut the output down to just the statistically salient OTUs. 


Unpack list of groups and pull out just 4 hour air filter dataset. 

```{r}
for (i in 1:length(names(groups))) {
  assign(names(groups)[i], unlist(groups[[i]]))
}

occ4f <- c(s14of, s24of, s34of, s14uf, s24uf, s34uf)
occ4f.table <- pb.1000[occ4f, ]
occ4f.table <- occ4f.table[, -which(colSums(occ4f.table) == 0)]
occ4f.map <- pb.map[occ4f, ]
```


Split this into two chunks - each takes a long time. 

```{r, cache=TRUE}
iv.location <- indval(occ4f.table, occ4f.map$location) 
```

```{r, cache=TRUE}
iv.person <- indval(occ4f.table, occ4f.map$occ.person2)
```

levels(occ4f.map$occ.person2)


Keep those that are significant before p-value adjustment. 

```{r}
p <- 0.05

these <- which(iv.location$pval <= p)
iv.location.table <- data.frame(iv.location$maxcls[these], 
  			  round(iv.location$indcls[these], 4), 
						iv.location$pval[these])
names(iv.location.table) <- c("cluster", "indicator_value", "probability")
iv.location.table$cluster[iv.location.table$cluster == '1'] <- 'occupied'
iv.location.table$cluster[iv.location.table$cluster == '2'] <- 'unoccupied'

these <- which(iv.person$pval <= p)
iv.person.table <- data.frame(iv.person$maxcls[these], 
				  round(iv.person$indcls[these], 4), 
						iv.person$pval[these])
names(iv.person.table) <- c("cluster", "indicator_value", "probability")
iv.person.table$cluster[iv.person.table$cluster == '1'] <- 'Subject1'
iv.person.table$cluster[iv.person.table$cluster == '2'] <- 'Subject2'
iv.person.table$cluster[iv.person.table$cluster == '3'] <- 'Subject3'
iv.person.table$cluster[iv.person.table$cluster == '4'] <- 'unoccupied'
```



Order by indicator value and tack on taxonomic information. 

```{r}
ivL <- iv.location.table[rev(order(iv.location.table$indicator_value)), ]
ivP <- iv.person.table[rev(order(iv.person.table$indicator_value)), ]

head(ivP)
ivL.names <- gsub('X', '', row.names(ivL))
ivL <- data.frame(cbind(ivL[, ], taxo[ivL.names, ]))

ivP.names <- gsub('X', '', row.names(ivP))
ivP <- cbind(ivP[, ], taxo[ivP.names, ])

ivL <- ivL[, c(2,3,10,1,9,8,7,6,5,4)]
ivP <- ivP[, c(2,3,10,1,9,8,7,6,5,4)]
```


Print out for later reference. 

```{r, results='asis'}
print(xtable(head(ivL, 20), digits=3), type='html')
print(xtable(head(ivP, 20), digits=3), type='html')
```


Clean up and save important bits to go to figures and tables script. 

```{r}
rm(list=ls()[which(ls() %in% names(groups))])
rm(occ4f, occ4f.table, occ4f.map, i, iv.location, iv.person, p, these, 
  iv.location.table, iv.person.table, ivL.names, ivP.names)
save(ivL, ivP, pb.order, taxo.tops, tops, consensus,
     file='../data/taxon_analysis.RData')
```



