# Personal Microbial Cloud

## Beta-diversity Data Analysis

Read in prepared workspace and load packages.

```{r}
load('../data/pb_data.RData') 
library(vegan)
library(labdsv)
library(ape)
library(xtable)
source('../data/pb_functions_for_manuscript.R')
```


Unpack list of groups.

```{r}
for (i in 1:length(names(groups))) {
  assign(names(groups)[i], unlist(groups[[i]]))
}
rm(i)
```


How do samples fall into groups? I.e., do we have balance in treatments?

```{r}
table(pb.map$person, pb.map$sampleType)
table(pb.map$location)
table(pb.map$duration)
table(pb.map$sampleType)
```



The canberra distance metric is used to calculate beta diversity. Then 2d NMDS objects are created for the whole experiment and for each treatment individually.

```{r, cache=TRUE}
pb.can <- vegdist(pb.1000, 'canberra')
```


Since most abundant taxa were removed as contaminants, try Bray-Curtis. 

```{r, cache=TRUE}
pb.bc <- vegdist(pb.1000, 'bray')
```





```{r, results='hide'}
# Filters
## 4h diff people
occ4f <- c(s24of, s34of, s14of, s24uf, s34uf, s14uf)
occ4f.table <- pb.1000[occ4f, ]
occ4f.map <- pb.map[occ4f, ]
occ4f.can <- as.dist(as.matrix(pb.can)[occ4f, occ4f]) 
occ4f.nmds <- bestnmds(occ4f.can)
occ <- which(occ4f.map$location == 'occ')
unocc <- which(occ4f.map$location == 'unocc')

## 2h diff people
occ2f <- c(s22of, s32of, s12of, s22uf, s32uf, s12uf)
occ2f.table <- pb.1000[occ2f, ]
occ2f.map <- pb.map[occ2f, ]
occ2f.can <- as.dist(as.matrix(pb.can)[occ2f, occ2f]) 
occ2f.nmds <- bestnmds(occ2f.can)
occ2 <- which(occ2f.map$location == 'occ')
unocc2 <- which(occ2f.map$location == 'unocc')

# Petri dishes
## 4h diff people
occ4p <- c(s24op, s34op, s14op, s24up, s34up, s14up)
occ4p.table <- pb.1000[occ4p, ]
occ4p.map <- pb.map[occ4p, ]
occ4p.can <- as.dist(as.matrix(pb.can)[occ4p, occ4p]) 
occ4p.nmds <- bestnmds(occ4p.can)
occ.p <- which(occ4p.map$location == 'occ')
unocc.p <- which(occ4p.map$location == 'unocc')

## 2h diff people
occ2p <- c(s22op, s32op, s12op, s22up, s32up, s12up)
occ2p.table <- pb.1000[occ2p, ]
occ2p.map <- pb.map[occ2p, ]
occ2p.can <- as.dist(as.matrix(pb.can)[occ2p, occ2p]) 
occ2p.nmds <- bestnmds(occ2p.can)
occ2.p <- which(occ2p.map$location == 'occ')
unocc2.p <- which(occ2p.map$location == 'unocc')
```


Pretty much all of these NMDS ordinations have high stress levels:

```{r, results='asis'}

nmds.stress <- data.frame(treatment = c('4hr filters', '2hr filters', '4hr petri dishes', '2hr petri dishes'), 
           stress = c(occ4f.nmds$stress, occ2f.nmds$stress, occ4p.nmds$stress, occ2p.nmds$stress),
           n = c(length(occ4f), length(occ2f), length(occ4p), length(occ2p)))
print(xtable(nmds.stress), type = "html")
```

But since statistics back up the visible patterns, it doesn't present much of a problem. 


The functions file (`pb_functions_for_mansucript.R`) contains a simple function to calculate richness, Shannon-Weiner diversity, and Evenness for each sample. 

```{r}
pb.hrj <- Evenness(pb.1000)

occ4f.map$R <- pb.hrj[occ4f, 'R']
occ4f.map$H1 <- pb.hrj[occ4f, 'H1']

occ2f.map$R <- pb.hrj[occ2f, 'R']
occ2f.map$H1 <- pb.hrj[occ2f, 'H1']

occ4p.map$R <- pb.hrj[occ4p, 'R']
occ4p.map$H1 <- pb.hrj[occ4p, 'H1']

occ2p.map$R <- pb.hrj[occ2p, 'R']
occ2p.map$H1 <- pb.hrj[occ2p, 'H1']
```


One last piece to assess the spread of points with beta-dispersion tests. It appears that s1 and s3 points are tightly grouped but s2 is not - this gives us a metric to report. 

```{r}
# 4of
pb.onlyocc.can <- as.dist(as.matrix(pb.can)[c(s24of, s34of, s14of), c(s24of, s34of, s14of)])
pb.onlyocc.map <- pb.map[c(s24of, s34of, s14of), ]
occ4f.bd <- betadisper(pb.onlyocc.can, pb.onlyocc.map$person)
occ4f.betad <- data.frame(dists = occ4f.bd$distances, person = pb.onlyocc.map$person)
occ4f.betad.aov <- anova(occ4f.bd)
#boxplot(occ4f.bd$distances~pb.onlyocc.map$person)

# 2of
pb.onlyocc2.can <- as.dist(as.matrix(pb.can)[c(s22of, s32of, s12of), c(s22of, s32of, s12of)])
pb.onlyocc2.map <- pb.map[c(s22of, s32of, s12of), ]
occ2f.bd <- betadisper(pb.onlyocc2.can, pb.onlyocc2.map$person)
occ2f.betad <- data.frame(dists = occ2f.bd$distances, person = pb.onlyocc2.map$person)
occ2f.betad.aov <- anova(occ2f.bd)
#boxplot(occ2f.bd$distances~pb.onlyocc2.map$person)

# 4op
pb.onlyoccp.can <- as.dist(as.matrix(pb.can)[c(s24op, s34op, s14op), c(s24op, s34op, s14op)])
pb.onlyoccp.map <- pb.map[c(s24op, s34op, s14op), ]
occ4p.bd <- betadisper(pb.onlyoccp.can, pb.onlyoccp.map$person)
occ4p.betad <- data.frame(dists = occ4p.bd$distances, person = pb.onlyoccp.map$person)
occ4p.betad.aov <- anova(occ4p.bd)
#boxplot(occ4p.bd$distances~pb.onlyoccp.map$person)

# 2of
pb.onlyocc2p.can <- as.dist(as.matrix(pb.can)[c(s22op, s32op, s12op), c(s22op, s32op, s12op)])
pb.onlyocc2p.map <- pb.map[c(s22op, s32op, s12op), ]
occ2p.bd <- betadisper(pb.onlyocc2p.can, pb.onlyocc2p.map$person)
occ2p.betad <- data.frame(dists = occ2p.bd$distances, person = pb.onlyocc2p.map$person)
occ2p.betad.aov <- anova(occ2p.bd)
#boxplot(occ2p.bd$distances~pb.onlyocc2p.map$person)

betad.aov.list <- list(occ4f.betad.aov, occ2f.betad.aov, occ4p.betad.aov, occ2p.betad.aov)
betad.table.list <- list(occ4f.betad, occ2f.betad, occ4p.betad, occ2p.betad)

rm(pb.onlyocc.can, pb.onlyocc.map, occ4f.bd,  
  pb.onlyocc2.can, pb.onlyocc2.map, occ2f.bd, 
  pb.onlyoccp.can, pb.onlyoccp.map, occ4p.bd,
  pb.onlyocc2p.can, pb.onlyocc2p.map, occ2p.bd, 
  occ4f.betad, occ4f.betad.aov, occ4p.betad, occ4p.betad.aov, 
  occ2f.betad, occ2f.betad.aov, occ2p.betad, occ2p.betad.aov)
```


After all of these data manipulations, some of these objects now go to a new script to make figures.  

```{r}
ls()
 
# clean groupings. 
rm(list=ls()[which(ls() %in% names(groups))])
rm(occ, unocc, occ.p, unocc.p, occ2, unocc2, occ2.p, unocc2.p, occ2f, occ2p, occ4f, occ4p)
ls()
# save the rest. 
save.image(file='../data/pb_analysis_to_figures.RData')
getwd()
```








