---
output: html_document
---


```{r config}
library(knitr)
opts_chunk$set(background='gray80', tidy=FALSE, cache=FALSE)
```

# PickleBox2014


## Analyze subsets



_James F Meadow_ (`jfmeadow at gmail dot com`)


```{r loadPackages, message=FALSE}
library(vegan)
library(labdsv)
library(ape)
library(scales)
```


Load particle count data and previous Rdata from `pb2014.Rmd` script. 

```{r importData}
rm(list=ls())
load('pb2014.RData')
load('particle/particleCounts.RData')
```

```{r setSeed}
set.seed(1314)
```

```{r keepOriginals}
tab.big <- tab
map.big <- map
taxa.big <- taxa
```


```{r getStats}
dim(tab.big)
row.names(tab.big)
```


Subset into individuals for exploration. 

```{r checkRarefactionLevels, eval=TRUE}
# sort(rowSums(tab.filtersHigh))
# sort(rowSums(tab.filtersLow))

s1high <- Reduce(intersect, list(filters, high, s1))
s2high <- Reduce(intersect, list(filters, high, s2))
s3high <- Reduce(intersect, list(filters, high, s3))
s4high <- Reduce(intersect, list(filters, high, s4))
s5high <- Reduce(intersect, list(filters, high, s5))
s6high <- Reduce(intersect, list(filters, high, s6))
s7high <- Reduce(intersect, list(filters, high, s7))
s8high <- Reduce(intersect, list(filters, high, s8))

s1low <- Reduce(intersect, list(filters, low, s1))
s2low <- Reduce(intersect, list(filters, low, s2))
s3low <- Reduce(intersect, list(filters, low, s3))
s4low <- Reduce(intersect, list(filters, low, s4))
s5low <- Reduce(intersect, list(filters, low, s5))
s6low <- Reduce(intersect, list(filters, low, s6))
s7low <- Reduce(intersect, list(filters, low, s7))
s8low <- Reduce(intersect, list(filters, low, s8))

allLow <- Reduce(intersect, list(filters, low, inside))
allHigh <- Reduce(intersect, list(filters, high, inside))
```

```{r configNMDS}
pbDist <- 'bray'
# pbDist <- 'canberra'
```



Change subject numbers so it flows with previous experiment. 

```{r newSubNames}
map$Subject2[map$Subject == 's01'] <- 's04'
map$Subject2[map$Subject == 's02'] <- 's05'
map$Subject2[map$Subject == 's03'] <- 's06'
map$Subject2[map$Subject == 's04'] <- 's07'
map$Subject2[map$Subject == 's05'] <- 's08'
map$Subject2[map$Subject == 's06'] <- 's09'
map$Subject2[map$Subject == 's07'] <- 's10'
map$Subject2[map$Subject == 's08'] <- 's11'
map$Subject2 <- factor(map$Subject2, 
                           levels=c('s11', 's09', 's08', 's05', 's07', 's04', 's10', 's06'))


```




Get sub dataset of occupied samples from low air flow. 


```{r getTopTaxa}
occLow <- which(map$Location == 'inside' & map$Treatment == 'low')
occTab <- tab[occLow, ]
occTab <- occTab[sort(row.names(occTab)), ]
occTab <- occTab/rowSums(occTab)
occMap <- map[row.names(occTab), ]
```

Create dataset of just targetted human-associated families. This list comes from the most influencial indicator taxa we found in the first experiment. 


```{r justHumanAssociated}
humAss <- sort(c(grep('Corynebacteriaceae', taxa$Family),
                 grep('Staphylococcaceae', taxa$Family),
                 grep('Streptococcaceae', taxa$Family),
                 grep('Lactobacillaceae', taxa$Family),
                 grep('Propionibacteriaceae', taxa$Family),
                 grep('Peptostreptococcaceae', taxa$Family),
                 grep('Bifidobacteriaceae', taxa$Family),
                 grep('Micrococcaceae', taxa$Family),
                 grep('Carnobacteriaceae', taxa$Family),
                 grep('Dietziaceae', taxa$Family), 
                 grep('Aerococcaceae', taxa$Family), 
                 grep('Tissierellaceae', taxa$Family)))


occTabHA <- occTab[, humAss]
occTaxHA <- taxa[humAss, ]

tabHA <- (tab/rowSums(tab))[, humAss]

identical(colnames(occTabHA), row.names(occTaxHA))
identical(row.names(occTabHA), row.names(occMap))
```


Check order of subjects and their colors. 

```{r checkLevels}
# occMap
unique(occMap$col)
levels(occMap$Subject)
```


Reorder subjects for plotting clarity. 

```{r plotAllHATogether}
occMap$SubOrder <- factor(occMap$Subject, levels=c('s08', 's06', 's05', 's02', 's04', 's01', 's07', 's03'))
subOrder <- c(8,6,5,2,4,1,7,3)

# Change names to make manuscript flow between 2 experiments. 
occMap$SubOrder2[occMap$SubOrder == 's01'] <- 's04'
occMap$SubOrder2[occMap$SubOrder == 's02'] <- 's05'
occMap$SubOrder2[occMap$SubOrder == 's03'] <- 's06'
occMap$SubOrder2[occMap$SubOrder == 's04'] <- 's07'
occMap$SubOrder2[occMap$SubOrder == 's05'] <- 's08'
occMap$SubOrder2[occMap$SubOrder == 's06'] <- 's09'
occMap$SubOrder2[occMap$SubOrder == 's07'] <- 's10'
occMap$SubOrder2[occMap$SubOrder == 's08'] <- 's11'
occMap$SubOrder2 <- factor(occMap$SubOrder2, 
                           levels=c('s11', 's09', 's08', 's05', 's07', 's04', 's10', 's06'))
```



Get a list of the subjects in order. 

```{r, fig.height=12, fig.width=8, eval=TRUE}
slows <- list(s1low, s2low, s3low, s4low, s5low, s6low, s7low, s8low)
cols <- unique(occMap$col)
```

Create function to plot the tree with the targeted OTU bar plots. 

```{r plotAllSubjectsTreeHABars} 
clMeth <- 'ward'
pbDist <- 'bray'
# set <- occLow
# rrare <- 0
# ptOffset <- .4

plotAllSubjectsTreeBars <- function(set, pbDist='bray', rrare=0, ptOffset=.4, 
                                    bar0=47, barlen=20, cutLine=FALSE, x..lim=65, keep=FALSE) {
  these <- set
  dat <- tabHA[these, ]
  dat <- dat[which(rowSums(dat) >= rrare), ]
  bars <- rowSums(dat)
  Dist <- vegdist(dat, pbDist)
  dend <- hclust(Dist, method=clMeth)
  phy <- as.phylo(dend)
  
  Map <- map[row.names(dat), ]
  phy$tip.label <- as.character(Map$Subject2)
  phy$order <- dend$order
  
  par(mar=c(4,0,1,1))
  plot(phy, show.tip.label=TRUE, use.edge.length=FALSE, 
    tip.color=Map$col, font=4, label.offset=1.5, cex=1.5, 
    edge.width=2, edge.color='gray50', x.lim=x..lim)
  tiplabels(pch=22, bg=Map$col, col='gray40', cex=2.4, adj=1.1)
  
  segments(rep(bar0, length(bars)), 1:length(bars), 
           rep(bar0, length(bars)) + (barlen*bars[dend$order]), 
           1:length(bars), 
           lwd=5, lend='butt', col=Map$col[dend$order])
  wlines <- c(bar0 + seq(.1, .6, .1)*barlen)
  segments(wlines, rep(0, length(wlines)), 
           wlines, rep(length(phy$tip.label)+1, length(wlines)), 
           col='white')
  axis(1, at=c(bar0, bar0+(barlen/2)), labels=c(0, .5))
  if(cutLine != FALSE) {
    cutt <- bar0 + (barlen*cutLine)
    abline(v=cutt, lty=3, col='gray60', lwd=2)
    axis(1, at=cutt, labels=cutLine)
    }
  if(keep == TRUE) {invisible(phy)}
  }
```




The subject that classify best generally have higher percentages of human associated bugs. 
Clearly, it works better for those who shed more during the experiment. 

This was for debugging - not evaluated. 

```{r bestSubjectsBoxplot, fig.width=5, fig.height=5, eval=FALSE}
# boxplot(rowSums(tabHA[occLow, ]) ~ map[occLow, 'Subject'], col=cols)
par(fg='gray30', col.axis='gray30')
boxplot(rowSums(occTabHA) ~ occMap$SubOrder2, 
        col=unique(occMap$col)[c(8,6,5,2,4,1,7,3)], 
        las=1, cex=0, border='gray40')
abline(h=.2, lty=3, col='gray60', lwd=2)
points(rowSums(occTabHA) ~ jitter(as.numeric(occMap$SubOrder2), .4), 
       pch=16, col=rgb(0,0,0,.5))
mtext('Relative Proportion of\nTargeted Human-Associated OTUs', 
      side=2, line=2.3)
mtext('Human Cloud Subjects', side=1, line=2.2)
```


Make a list of subjects who clustered well for further analysis. 


```{r makeBestSubs}
bestSubs <- c(Reduce(intersect, list(occLow, c(s8, s6, s5, s4, s2))))
```


Function for plotting individual occupant trees with their exhaust and supply samples. 

```{r plotEachSubjectTree} 
plotEachSubjectTree <- function(lst, pbDist, rrare=0, ptOffset=.4, 
                                bar0=rep(16, length(lst)), 
                                barlen=rep(4, length(lst)), 
                                cutLine=FALSE) {
  for(i in 1:length(lst)) {
    these <- lst[[i]]
    dat <- tabHA[these, ]
    dat <- dat[which(rowSums(dat) >= rrare), ]
    bars <- rowSums(dat)
    Dist <- vegdist(dat, pbDist)
    dend <- hclust(Dist, method=clMeth)
    phy <- as.phylo(dend)
#     phy$tip.label <- gsub('^[[:alnum:]]{1}\\.[[:alnum:]]{1}\\.', '', phy$tip.label)
#     phy$tip.label <- gsub('\\.[[:digit:]]{1}$', '', phy$tip.label)
#     phy$tip.label <- gsub('s[[:digit:]]{2}\\.sup[[:digit:]]{1}$', 'sup', phy$tip.label)
#     phy$tip.label <- gsub('s[[:digit:]]{2}\\.exh[[:digit:]]{1}$', 'exh', phy$tip.label)

    bari <- bar0[i]
    if (length(barlen) == 1) {barleni <- barlen}
    if (length(barlen) !=1) {barleni <- barlen[i]}
    Map <- map[row.names(dat), ]
    phy$tip.label <- as.character(Map$Subject2)
    
    par(mar=c(2, 0, 1, 1))
    plot(phy, show.tip.label=FALSE, use.edge.length=FALSE, 
      #tip.color=Map$col, font=4, label.offset=1.5, 
      edge.width=2, edge.color='gray50', x.lim=bari+(barleni*.6))
    tiplabels(pch=22, bg=Map$col, col='gray40', cex=1.6, adj=1.1, lwd=.3)    

    segments(rep(bari, length(bars)), 1:length(bars), 
             rep(bari, length(bars)) + (barleni*bars[dend$order]), 
             1:length(bars), 
             lwd=5, lend='butt', col=Map$col[dend$order])
    wlines <- c(bari + seq(.1, .6, .1)*barleni)
    segments(wlines, rep(0, length(wlines)), 
             wlines, rep(length(phy$tip.label)+1, length(wlines)), 
             col='white')
    axis(1, at=c(bari, bari+(barleni/2)), labels=c(0, .5), col='gray30')
    axis(1, at=wlines[1:5], labels=FALSE, tck=-.02, col='gray30')
    if(cutLine != FALSE) {
      cutt <- bari + (barleni*cutLine)
      abline(v=cutt, lty=3, col='gray60', lwd=2)
      axis(1, at=cutt, labels=cutLine)
      }      
    }
  }
```

```{r setupGroups}
lowList <- list(s1low, s2low, s3low, s4low, s5low, s6low, s7low, s8low)
highList <- list(s1high, s2high, s3high, s4high, s5high, s6high, s7high, s8high)
```


--------------


# barplots


Set up data frames to subset distance matrices into barplots. 

```{r getLowSelfDistBarplot}
se <- function(x) {sd(x)/sqrt(length(x))}
distLowBarDF <- data.frame(matrix(0, nrow=8, ncol=6))
distLowBarDFSE <- data.frame(matrix(0, nrow=8, ncol=6))
names(distLowBarDF) <- c('self', 'selfExh', 'selfSup', 'other', 'otherExh', 'otherSup')
names(distLowBarDFSE) <- c('self', 'selfExh', 'selfSup', 'other', 'otherExh', 'otherSup')
pbDist <- 'jaccard'


for(i in 1:8) {
  these <- lowList[[i]]
  dat <- tab[these, ]
  dat <- dat[which(rowSums(dat) >=rrare), ]
  dat <- dat[sort(row.names(dat)), ]
  Dist <- 1-vegdist(dat, pbDist)
  Map <- map[row.names(dat), ]
  
  self <- which(Map$Location == 'inside')
  exh <- which(Map$Location == 'exh')
  sup <- which(Map$Location == 'sup')
  
  DistSelf <- as.dist(as.matrix(Dist)[self, self])
  DistExh <- as.matrix(Dist)[self, exh]
  DistSup <- as.matrix(Dist)[self, sup]
  
  distLowBarDF$self[i] <- mean(DistSelf)
  distLowBarDF$selfExh[i] <- mean(DistExh)
  distLowBarDF$selfSup[i] <- mean(DistSup)
  
  distLowBarDFSE$self[i] <- se(DistSelf)
  distLowBarDFSE$selfExh[i] <- se(DistExh)
  distLowBarDFSE$selfSup[i] <- se(DistSup)
  
  }
  
```

```{r getHighSelfDistBarplot}
distHighBarDF <- data.frame(matrix(0, nrow=8, ncol=6))
distHighBarDFSE <- data.frame(matrix(0, nrow=8, ncol=6))
names(distHighBarDF) <- c('self', 'selfExh', 'selfSup', 'other', 'otherExh', 'otherSup')
names(distHighBarDFSE) <- c('self', 'selfExh', 'selfSup', 'other', 'otherExh', 'otherSup')

for(i in 1:8) {
  these <- highList[[i]]
  dat <- tab[these, ]
  dat <- dat[which(rowSums(dat) >=rrare), ]
  dat <- dat[sort(row.names(dat)), ]
  Dist <- 1-vegdist(dat, pbDist)
  Map <- map[row.names(dat), ]
  
  self <- which(Map$Location == 'inside')
  exh <- which(Map$Location == 'exh')
  sup <- which(Map$Location == 'sup')
  
  DistSelf <- as.dist(as.matrix(Dist)[self, self])
  DistExh <- as.matrix(Dist)[self, exh]
  DistSup <- as.matrix(Dist)[self, sup]
  
  distHighBarDF$self[i] <- mean(DistSelf)
  distHighBarDF$selfExh[i] <- mean(DistExh)
  distHighBarDF$selfSup[i] <- mean(DistSup)

  distHighBarDFSE$self[i] <- se(DistSelf)
  distHighBarDFSE$selfExh[i] <- se(DistExh)
  distHighBarDFSE$selfSup[i] <- se(DistSup)
  }
  
```

```{r getLowOtherBarplot}
distLowBarDF; distLowBarDFSE
subs <- as.character(subs)
dat <- tab[which(map$Treatment == 'low'), ]
dat <- dat[which(rowSums(dat) >= rrare), ]
Map <- map[row.names(dat), ]
Dist <- as.matrix(1-vegdist(dat, pbDist))

for(i in 1:8) {
  distLowBarDF$other[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'inside'])  
  distLowBarDF$otherExh[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'exh'])  
  distLowBarDF$otherSup[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'sup'])  

  distLowBarDFSE$other[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'inside'])  
  distLowBarDFSE$otherExh[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'exh'])  
  distLowBarDFSE$otherSup[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'sup'])  

  }
```

```{r getHighOtherBarplot}
distHighBarDF$selfExh[which(distHighBarDF$selfExh == 'NaN')] <- NA
distHighBarDFSE$selfExh[which(distHighBarDFSE$selfExh == 'NaN')] <- NA
subs <- as.character(subs)
dat <- tab[which(map$Treatment == 'high'), ]
dat <- dat[which(rowSums(dat) >= rrare), ]
Map <- map[row.names(dat), ]
Dist <- as.matrix(1-vegdist(dat, pbDist))

for(i in 1:8) {
  distHighBarDF$other[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'inside'])  
  distHighBarDF$otherExh[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'exh'])  
  distHighBarDF$otherSup[i] <- 
    mean(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'sup'])  

  distHighBarDFSE$other[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'inside'])  
  distHighBarDFSE$otherExh[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'exh'])  
  distHighBarDFSE$otherSup[i] <- 
    se(Dist[Map$Subject == subs[i] & Map$Location == 'inside', 
              Map$Subject != subs[i] & Map$Location == 'sup'])  
  }
```


Put all of that into smaller plotting dataframes. 

```{r plottingDataFrames}
se. <- function(x) {
  len <- length(x[which(!is.na(x))])
  sd(x, na.rm=TRUE)/sqrt(len)
  }

plotLowDF <- data.frame(mean = apply(distLowBarDF, 2, mean), 
                        se = apply(distLowBarDF, 2, se.))
plotHighDF <- data.frame(mean = apply(distHighBarDF, 2, mean, na.rm=TRUE), 
                        se = apply(distHighBarDF, 2, se.))
plotLowDF <- plotLowDF[c(1, 4, 2, 5, 3, 6), ]
plotHighDF <- plotHighDF[c(1, 4, 2, 5, 3, 6), ]
```

```{r setupBarPlot, eval=TRUE}
space=rep(c(.4, .1), 3)
cols <- c('gray', 'gray50')
ylims <- c(0, .25)
```

```{r makeLowBarPlot, fig.height=8, fig.width=6, eval=TRUE}
df <- plotLowDF
len <- 0.03
# pdf('sixLowBar.pdf', width=6, height=8)
par(mar=c(5, 5, 4, 1), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col=cols, border='white', space=space, 
                ylim=ylims, ylab='Canberra Similarity')
abline(h=seq(0, .4, .05), col='white')
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('inside', 'exhaust', 'supply'), side=1, line=0.5, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1.4, col='gray30')
legend('topright', legend=c('self vs. self', 'self vs. other'), pt.cex=3, 
       pch=15, col=c('gray', 'gray50'), bty='n', cex=1.4, text.col='gray30')           
mtext('1 Air Change / Hour', col='gray30', line=1, font=2, cex=2)
# dev.off()      
```

```{r makeHighBarPlot, fig.height=8, fig.width=6, eval=TRUE}
df <- plotHighDF
len <- 0.03
# pdf('sixBarHigh.pdf', width=6, height=8)
par(mar=c(5, 4, 4, 1), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col=cols, border='white', space=space, 
                ylim=ylims)
abline(h=seq(0, .4, .05), col='white')
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('inside', 'exhaust', 'supply'), side=1, line=0.5, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1.4, col='gray30')
legend('topright', legend=c('self vs. self', 'self vs. other'), pt.cex=3, 
       pch=15, col=c('gray', 'gray50'), bty='n', cex=1.4, text.col='gray30')           
mtext('3 Air Change / Hour', col='gray30', line=1, font=2, cex=2)
# dev.off()      
```


Make supplementary barplot. 

```{r makeLowHighBarPlot, fig.height=5, fig.width=7, eval=TRUE}
df <- plotLowDF
len <- 0.03
space <- rep(c(.8, .02), 3)
pdf('suppBarplots2.pdf', width=7, height=5)
layout(matrix(c(1,2), 1, 2), widths=c(1, .9))
par(mar=c(5, 5, 4, 1), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col='gray', border='white', space=space, 
                ylim=ylims, ylab='Jaccard Similarity', yaxt='n')
abline(h=seq(0, .4, .05), col='white')
axis(2, tck=.02, labels=FALSE)
axis(2, tick=FALSE, line=-.6)
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('self', 'other'), side=1, line=0, 
      at=mids, font=1, cex=.8, col='gray30')
mtext(c('Inside', 'Exhaust', 'Supply'), side=1, line=1, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1, col='gray30')
mtext('1 Air Change / Hour', col='gray30', line=1, font=2, cex=1)
# legend('topright', legend=c('self vs. self', 'self vs. other'), pt.cex=3, 
#        pch=15, col=c('gray', 'gray50'), bty='n', cex=1.4, text.col='gray30')           

df <- plotHighDF
par(mar=c(5, 3, 4, 1), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col='gray', border='white', space=space, 
                ylim=ylims, yaxt='n')
abline(h=seq(0, .4, .05), col='white')
axis(2, tck=.02, labels=FALSE)
axis(2, tick=FALSE, line=-.6)
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('self', 'other'), side=1, line=0, 
      at=mids, font=1, cex=.8, col='gray30')
mtext(c('Inside', 'Exhaust', 'Supply'), side=1, line=1, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1, col='gray30')
mtext('3 Air Changes / Hour', col='gray30', line=1, font=2, cex=1)
dev.off()      

```




--------------------------------------

## Particle count data

```{r}
boxplot(Agg.df$diff10 ~ Agg.df$group, cex=.6, col=unique(occMap$col), 
        pch=16, las=1, xaxt='n', ylab='')
mtext(expression(paste('Particles (>10', mu, 'm) L' ^{-1},  ' min' ^{-1})), 
	side=2, line=2)
abline(h=0, col='gray', lty=3, lwd=2)
mtext('People', side=1)

```


```{r aggregateParticleData}
partMean <- aggregate(Agg.df[, c('diff10', 'diff510', 'diff35', 'diff13')], by=list(Agg.df$group), FUN='mean')
partSD <- aggregate(Agg.df[, c('diff10', 'diff510', 'diff35', 'diff13')], by=list(Agg.df$group), FUN='sd')
partSE <- aggregate(Agg.df[, c('diff10', 'diff510', 'diff35', 'diff13')], by=list(Agg.df$group), FUN='se')
```


```{r aggregateHASums}
haMean <- aggregate(rowSums(occTabHA), by=list(occMap$Subject), FUN='mean')
names(haMean) <- c('group', 'mean')
haMean$sd <- aggregate(rowSums(occTabHA), by=list(occMap$Subject), FUN='sd')$x
haMean$se <- aggregate(rowSums(occTabHA), by=list(occMap$Subject), FUN='se')$x
```



Set up the plot, with three sizes of air particles on the same plot. 


```{r plotHAandParticles10}
# xmean <- partMean$diff10
# ymean <- haMean$mean
# xse <- partSE$diff10
# yse <- haMean$se
plotPartHA <- function(xmean=xmean, ymean=ymean, xse=xse, yse=yse, pch=21, New=FALSE) {
  x.lim <- c(min(xmean - xse), 
             max(xmean + xse))
  y.lim <- c(min(ymean - yse), 
             max(ymean + yse))  
  if (New) {
    par(new=TRUE)
    plot(xmean, ymean, type='n', 
         xlim=x.lim, ylim=y.lim, las=1, 
         ann=FALSE, axes=FALSE, bty='n')
    }
  else {
    par(new=FALSE)
    plot(xmean, ymean, type='n', 
         xlim=x.lim, ylim=y.lim, las=1, 
         ann=FALSE, xaxt='n', bty='n')
    }
  arrows(xmean, ymean + yse, 
         xmean, ymean - yse, 
         code=3, angle=90, length=.02)
  arrows(xmean + xse, ymean, 
         xmean - xse, ymean, 
         code=3, angle=90, length=.02)
  points(xmean, ymean, 
         bg=alpha(unique(occMap$col), alpha=.7), 
         cex=3, pch=pch)
}  


plotPartHA(partMean$diff10, haMean$mean, partSE$diff10, haMean$se)
plotPartHA(partMean$diff510, haMean$mean, partSE$diff510, haMean$se, New=TRUE, pch=22)
plotPartHA(partMean$diff35, haMean$mean, partSE$diff35, haMean$se, New=TRUE, pch=24)
plotPartHA(partMean$diff13, haMean$mean, partSE$diff13, haMean$se, New=TRUE, pch=23)

plotPartHA(partMean$diff10, haMean$mean, partSE$diff10, haMean$se)
plotPartHA(partMean$diff510, haMean$mean, partSE$diff510, haMean$se, New=TRUE, pch=22)
plotPartHA(partMean$diff35, haMean$mean, partSE$diff35, haMean$se, New=TRUE, pch=24)
# plotPartHA(partMean$diff13, haMean$mean, partSE$diff13, haMean$se, New=TRUE, pch=23)


cor.test(partMean$diff10, haMean$mean)$estimate
cor.test(partMean$diff510, haMean$mean)$estimate
cor.test(partMean$diff35, haMean$mean)$estimate
cor.test(partMean$diff13, haMean$mean)$estimate
```



A version of the tree plotting function that keeps the order of tips, so that data can be put on the tips in the right order.

```{r plotAllSubjectsTreeHABars2} 
clMeth <- 'ward'
pbDist <- 'bray'

plotAllSubjectsTreeBars2 <- function(set, pbDist='bray', rrare=0, ptOffset=2, 
                                    bar0=47, barlen=20, cutLine=FALSE, x..lim=65, keep=FALSE) {
  these <- set
  dat <- tabHA[these, ]
  dat <- dat[which(rowSums(dat) >= rrare), ]
  bars <- rowSums(dat)
  Dist <- vegdist(dat, pbDist)
  dend <- hclust(Dist, method=clMeth)
  phy <- as.phylo(dend)
  
  Map <- map[row.names(dat), ]
  phy$tip.label <- as.character(Map$Subject2)
  phy$order <- dend$order
  
  if (keep == FALSE) {
    par(mar=c(5,0,1,1))
    plot(phy, show.tip.label=FALSE, use.edge.length=FALSE, 
      tip.color=Map$col, font=4, label.offset=1.5, cex=1.5, 
      edge.width=2, edge.color='gray50', x.lim=x..lim)
    tiplabels(pch=15, col=Map$col, cex=2, adj=ptOffset)
    }
  
  if(keep == TRUE) {invisible(phy)}
  }
```



----------------------------

```{r changepbdist}
pbDist <- 'bray'
```



## Check signif at splitting nodes. 

In order to put statistics on clade separation, this was set up to get adonis test based on the current tree. After running these, statistics were put manually on the trees of the final version of the manuscript figures. 

```{r}
Phy <- plotAllSubjectsTreeBars2(set = bestSubs, pbDist = 'bray', rrare = 0, keep = TRUE)
Phy$order
Phy2 <- plotAllSubjectsTreeBars2(set=occLow, pbDist='bray', rrare=.2, keep=TRUE)
# row.names(tabHA)[occLow]
row.names(tabHA)[bestSubs][Phy$order]

# testThese is the order on the tree being tested, from bottom to top. 
checkSig <- function(these=bestSubs, order=Phy$order, testThese, pbDist='bray', clusters) {
  TAB <- tabHA[these, ][order, ][testThese, ]
  DIST <- vegdist(TAB, pbDist)
  print(adonis(DIST ~ clusters))
  }
```


Test clades in the bottom tree (bestSubs)

```{r}
# all in bottom tree, broken in at root. 
checkSig(testThese=c(1:25), clusters=c(rep('a', 8), rep('b', 17)))

# bottom two clades
checkSig(testThese=c(1:8), clusters=c(rep('a', 3), rep('b', 5)))

# top 3 clades, broken at 13/14
checkSig(testThese=c(9:25), clusters=c(rep('a', 5), rep('b', 12)))

# top 2 clades, red vs orange
checkSig(testThese=c(14:25), clusters=c(rep('a', 4), rep('b', 8)))

# Check top clade mess
checkSig(testThese=c(18:25), clusters=c(rep('a', 3), rep('b', 5)))
```


Test clades in top tree (`rrare=.2`)

```{r}
check2 <- which(rowSums(tabHA)[occLow] > .2)
row.names(tabHA)[occLow][check2][Phy2$order]
these.2 <- occLow[check2]
# bottom two clades
checkSig(these=these.2, order=Phy2$order, testThese=c(1:9), clusters=c(rep('a', 5), rep('b', 4)))

# s08 vs all other above
checkSig(these=these.2, order=Phy2$order, testThese=c(6:22), clusters=c(rep('a', 4), rep('b', 13)))

# s05 vs all above
checkSig(these=occLow[check2], order=Phy2$order, testThese=c(10:22), clusters=c(rep('a', 4), rep('b', 9)))

# s09 vs all above
checkSig(these=occLow[check2], order=Phy2$order, testThese=c(14:22), clusters=c(rep('a', 4), rep('b', 5)))

# s07 vs s04
checkSig(these=occLow[check2], order=Phy2$order, testThese=c(18:22), clusters=c(rep('a', 3), rep('b', 2)))
```



---------------------------------------


## Big figure

This is the figure that ends up in the manuscript. 

```{r plotBigFigure, fig.width=11, fig.height=7}
pdf('secondExpPanels.pdf', width=11, height=7, useDingbats=FALSE)
layout(matrix(c(1,1,2:5), 2, 3), widths=c(1, .7, .7), heights=c(.9, 1))

plotAllSubjectsTreeBars(occLow, pbDist=pbDist, rrare=0, cutLine=.2, 
                        x..lim=72, bar0=48, barlen=26)
par(xpd=TRUE)
text(52, -3, 'Targeted OTUs', col='gray20', cex=1.2) 
par(xpd=FALSE)

par(fg='gray30', col.axis='gray30', bty='l')
boxplot(rowSums(occTabHA) ~ occMap$SubOrder2, 
        col=unique(occMap$col)[c(8,6,5,2,4,1,7,3)], 
        las=1, cex=0, border='gray40')
abline(h=.2, lty=3, col='gray60', lwd=2)
points(rowSums(occTabHA) ~ jitter(as.numeric(occMap$SubOrder2), .4), 
       pch=16, col=rgb(0,0,0,.5))
mtext('Relative Abundance of\nTargeted Human-Associated OTUs', 
      side=2, line=2.3)
mtext('Human Cloud Subjects', side=1, line=2.2)

par(mar=c(8, 0, 1, 5))
plotPartHA(partMean$diff10, haMean$mean, partSE$diff10, haMean$se)
axis(side=1, line=0, tck=0.01, labels=FALSE)
axis(side=1, line=-1, tck=0.0, lwd=0, lwd.ticks=0)
plotPartHA(partMean$diff510, haMean$mean, partSE$diff510, haMean$se, New=TRUE, pch=22)
axis(side=1, line=2, tck=0.01, labels=FALSE)
axis(side=1, line=1, tck=0.0, lwd=0, lwd.ticks=0)
plotPartHA(partMean$diff35, haMean$mean, partSE$diff35, haMean$se, New=TRUE, pch=24)
axis(side=1, line=4, tck=0.01, labels=FALSE)
axis(side=1, line=3, tck=0.0, lwd=0, lwd.ticks=0)
# plotPartHA(partMean$diff13, haMean$mean, partSE$diff13, haMean$se, New=TRUE, pch=23)
# axis(side=1, line=6, tck=0.01, labels=FALSE)
# axis(side=1, line=5, tck=0.0, lwd=0, lwd.ticks=0)

abline(h=.2, lty=3, col='gray60', lwd=2)
par(xpd=TRUE)
points(rep(146, 3), c(.05, 0, -.05), 
       pch=c(21, 22, 24), cex=2, 
       bg='gray40', col='gray40')
text(rep(147, 3), c(.05, 0, -.05)-.005, 
     c(expression(paste('>10', mu, 'm')), 
       expression(paste('5-10', mu, 'm')), 
       expression(paste('2.5-5', mu, 'm'))), pos=4)
mtext(expression(paste('Airborne Particle Load L' ^{-1},  ' 10-min' ^{-1})), 
      side=1, line=6)
mtext('Relative Abundance of\nTargeted Human-Associated OTUs', 
      side=2, line=2.3)


par(xpd=FALSE)
plotAllSubjectsTreeBars(occLow, pbDist=pbDist, rrare=.2, 
                        bar0=28, barlen=25, x..lim=45)
par(xpd=TRUE)
text(34, -3, 'Targeted OTUs', col='gray20', cex=1.2) 
par(xpd=FALSE)


plotAllSubjectsTreeBars(bestSubs, pbDist=pbDist, rrare=0, 
                        bar0=32, barlen=27, x..lim=50)
par(xpd=TRUE)
text(39, -3, 'Targeted OTUs', col='gray20', cex=1.2) 
par(xpd=FALSE)


par(mfrow=c(1,1), mar=c(0,0,0,0), new=TRUE)
plot(1,1, xlim=c(0, 1), ylim=c(0, 1), xaxs='i', yaxs='i', 
     axes=FALSE, ann=FALSE, type='n')




text(c(.05, .67, .67, .75, .75), 
     c(.95, .95, .47, .95, .47), 
     c('(a)', '(b)', '(c)', '(d)', '(e)'),
     col='gray20', font=2, cex=1.5)

dev.off()
```


Manuscript figure that shows three example trees - each with occupied, exhaust, and supply samples. 

```{r plot3trees}
pdf('threeExampleTrees.pdf', width=2, height=4.5)
# par(mfrow=c(3, 1))
layout(matrix(c(1, 2, 3, 4), 4, 1), heights=c(1,1,1,.2))
plotEachSubjectTree(lst=lowList[c(8,2,7)], pbDist=pbDist, 
                    bar0=c(11.2, 10.2, 12.3), barlen=c(8.4, 7.6, 9.3))




par(mfrow=c(1,1), mar=c(0,0,0,0), new=TRUE)
plot(1,1, xlim=c(0, 1), ylim=c(0, 1), xaxs='i', yaxs='i', 
     axes=FALSE, ann=FALSE, type='n')
text(.75, .04, 'Relative Abundance\nof Targeted OTUs', 
     col='gray20', cex=.6)
text(c(rep(0, 3)), 
     c(.96, .66, .34), 
     c('(a)', '(b)', '(c)'),
     col='gray20', font=2, cex=1, pos=4)

legdf <- data.frame(x=c(.04, .04, .04, .1, .16), 
                    y=c(.03, .06, .09, .09, .09), 
                    col=c('#ECCBAE', # exhaust
                          'gray', # supply
                          '#ABDDDE', # s08
                          '#FF0000', # s02
                          '#D69C4E'  # s07
                          ))  
                          
points(legdf$x, legdf$y, pch=22, col='gray30', 
       bg=as.character(legdf$col), cex=1.3, lwd=.3)
text(c(.16, .04, .04), c(.09, .06, .03), 
     c('  Subjects', '.......... Supply Air', '.......... Exhaust Air'), 
     pos=4, col='gray30', cex=.6)


dev.off()


# (lst, pbDist, rrare=0, ptOffset=.4, 
# bar0=rep(16, length(lst)), barlen=4, cutLine=FALSE) 

```





These chunks of code are not 

```{r setupBarPlotX}
space=rep(c(.4, .1), 3)
cols <- c('gray', 'gray50')
```


```{r setIndividualBarOffsetX}
lowLim <- c(16, 14, 14, 15, 16, 14, 16, 15)
highLim <- c(15, 14, 16, 15, 14, 16, 16, 16)
```


```{r bigPlotSmallTrees, fig.height=14, fig.width=6, eval=FALSE}
pdf('bigPlotManyTrees.pdf', height=14, width=6)
layout(matrix(c(1, 3:10, 1, 11:18, 2, 19:26, 2, 27:34), 9, 4), 
       heights=c(3, rep(1, 8)))

df <- plotLowDF
len <- 0.03
# pdf('sixBarBoth.pdf', width=10, height=8)
# layout(matrix(c(1,2), 1, 2), widths=c(1, .8))
par(mar=c(5, 5, 3, 1), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col=cols, border='white', space=space, 
                ylim=c(0, .4), ylab='Canberra Similarity')
abline(h=seq(0, .4, .05), col='white')
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('inside', 'exhaust', 'supply'), side=1, line=0.5, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1, col='gray30')
mtext('1 Air Change / Hour', col='gray30', line=.5, font=2, cex=1.2)

df <- plotHighDF
par(mar=c(5, 2, 3, 4), las=1)
mids <- barplot(as.matrix(df$mean), beside=TRUE, 
                col=cols, border='white', space=space, 
                ylim=c(0, .4), yaxt='n')
abline(h=seq(0, .4, .05), col='white')
arrows(mids, df$mean+df$se, mids, df$mean-df$se, 
       code=3, length=len, angle=90, col='gray30', lwd=2)
par(xpd=TRUE)
segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
mtext(c('inside', 'exhaust', 'supply'), side=1, line=0.5, 
      at=c(mean(mids[1:2]), mean(mids[3:4]), mean(mids[5:6])), 
      font=2, cex=1, col='gray30')
legend('topright', legend=c('self vs. self', 'self vs. other'), pt.cex=3, 
       pch=15, col=c('gray', 'gray50'), bty='n', cex=1.4, text.col='gray30')           
mtext('3 ACH', col='gray30', line=.5, font=2, cex=1.2)


plotEachSubjectTree(lowList[subOrder], pbDist, rrare=0, ptOffset=33, 
                    bar0=lowLim[subOrder]-3, barlen=15)
plotEachSubjectTree(highList[subOrder], pbDist, rrare=0, ptOffset=33, 
                    bar0=highLim[subOrder]-3, barlen=15)


par(mar=c(0.2,1,1,1))
space=rep(c(.4, .1), 3)
cols <- c('gray', 'gray50')

df <- distLowBarDF[subOrder, c(1,4,2,5,3,6)]
dfSE <- distLowBarDFSE[subOrder, c(1,4,2,5,3,6)]
barcols <- unique(occMap$col)[subOrder]

for(i in 1:8) {
  mids <- barplot(as.matrix(df[i, ]), beside=TRUE, 
                  col=c(barcols[i], cols[2]), border='white', 
                  space=space, ylim=c(0, .5), ylab='', 
                  axes=FALSE, names.arg=rep('', 6))
  abline(h=seq(0, .5, .1), col='white')
  arrHigh <- unlist(df[i,]+dfSE[i,])
  arrLow <- unlist(df[i,]-dfSE[i,])
  arrows(mids, arrHigh, mids, arrLow, 
         code=3, length=len, angle=90, col='gray30', lwd=2)
  par(xpd=TRUE)
  segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
}

df <- distHighBarDF[subOrder, c(1,4,2,5,3,6)]
dfSE <- distHighBarDFSE[subOrder, c(1,4,2,5,3,6)]
barcols <- unique(occMap$col)[subOrder]

for(i in 1:8) {
  mids <- barplot(as.matrix(df[i, ]), beside=TRUE, 
                  col=c(barcols[i], cols[2]), border='white', 
                  space=space, ylim=c(0, .5), ylab='', 
                  axes=FALSE, names.arg=rep('', 6))
  abline(h=seq(0, .5, .1), col='white')
  arrHigh <- unlist(df[i,]+dfSE[i,])
  arrLow <- unlist(df[i,]-dfSE[i,])
  arrows(mids, arrHigh, mids, arrLow, 
         code=3, length=len, angle=90, col='gray30', lwd=2)
  par(xpd=TRUE)
  segments(mids[c(1, 3, 5)]-.5, 0, mids[c(2, 4, 6)]+.5, 0, lwd=2, col='gray40')
}
dev.off()
```












----------

This creates the figure with a tree, and indicator taxa for each clade. Several steps to set it up. 


```{r treeAnalysis}
# these <- c(1, 2, 4, 5, 7, 8, 10, 15, 19, 24, 28, 31, 46, 48)
these2 <- c(1, 2, 5, 10, 7, 8, 15, 19, 24, 28, 31, 48)
these3 <- c(7, 5, 11, 6, 9, 10, 8, 12)
## get tip order from the tree plotting function
Phy <- plotAllSubjectsTreeBars2(bestSubs, pbDist=pbDist, rrare=0, keep=TRUE)

names(Phy)
tipOrderNames <- Phy$tip.label[Phy$order]
tipOrderRN <- row.names(tabHA)[bestSubs][Phy$order]

treeBars <- tabHA[tipOrderRN, these2]
treeCols <- map[tipOrderRN, 'col']
treeTaxa <- taxa[colnames(treeBars), ]
```


```{r}
checkRelAbu <- colnames(treeBars)[these3]
mapTreeBars <- map[row.names(treeBars), ]
subsTreeBars <- unique(mapTreeBars$Subject2)
for(i in 1:length(subsTreeBars)) {
  these <- which(mapTreeBars$Subject2 == subsTreeBars[i])
  mapThese <- mapTreeBars[these, ]
  newTab <- tabHA[row.names(mapThese), ]
  newSums <- sort(colSums(newTab), decreasing=TRUE)
  print(subsTreeBars[i])
  print(names(newSums)[1:10])
  }

```


```{r checkIndicators}
# Some are selected as indicators for a single person. 
# This is the same order as the following figure. 

checkInd <- colnames(treeBars)[these3]
indBars <- indval(treeBars[, checkInd], clustering=treeCols)
indBars  # all are significant indicators for their respective occupant
```




Create function to add barplot to the base tree. 

```{r barsFunction}
addBars <- function(thisone=1, bar0=30, bar1=39, Lines=FALSE, labcol='gray30', 
                    text1='', text2='') {
  bars <- treeBars[, thisone]/max(treeBars[, thisone])
  barlen <- bar1-bar0
  lab <- which(bars == 1)
  labMax <- treeBars[lab, thisone]
  labTotal <- format(labMax, digits=2)
  
  if(Lines) {
    checkLines <- seq(0, 150, 5)
    abline(v=checkLines)
    mtext(checkLines, at=checkLines, cex=.4, line=0)
    }

  segments(rep(bar0, length(bars)), 1:length(bars), 
           rep(bar0, length(bars)) + (barlen*bars), 1:length(bars), 
           lwd=13, lend='butt', col=treeCols)
  abline(v=bar0, lwd=1, col='gray70')
  text(bar0-1, lab-.05, labTotal, pos=4, font=2, col=labcol, cex=.7)
  par(xpd=TRUE)
#   text(bar0+5, -.5, treeTaxa[thisone, 'Genus'], 
#        srt=30, pos=2, cex=.7, font=4)
  text(bar0+5, -.5, text1, srt=35, pos=2, cex=.6, font=4)
  text(bar0+6, -1.1, text2, srt=35, pos=2, cex=.5, font=1)
  par(xpd=FALSE)
  }
```


This makes the tree, and the addBars function puts the barplots on one by one. 

```{r treeFigWithIndicBars}
pdf('clusterTaxa.pdf', width=5, height=5, useDingbats=FALSE)
plotAllSubjectsTreeBars2(bestSubs, pbDist=pbDist, rrare=0, ptOffset=3, 
                        bar0=32, barlen=27, x..lim=110, keep=FALSE)

addBars(thisone=7, bar0=30, bar1=39, 
        text1='Lactobacillus crispatus', 
        text2='NR_119274.1 (100% similar)')
addBars(thisone=5, bar0=40, bar1=49, 
        text1='Gardnerella vaginalis', 
        text2='NR_074227.1 (100% similar)')
addBars(thisone=11, bar0=50, bar1=59, 
        text1='Corynebacterium massiliense', 
        text2='NR_044182.1 (100% similar)')
addBars(thisone=6, bar0=60, bar1=69, 
        text1='Corynebacterium imitans', 
        text2='NR_026369.1 (100% similar)')
addBars(thisone=9, bar0=70, bar1=79, 
        text1='Corynebacterium epidermidicanis', 
        text2='NR_108533.1 (98% similar)')
addBars(thisone=10, bar0=80, bar1=89, labcol='gray90', 
        text1='Anaerococcus octavius', 
        text2='NR_026360.1 (97% similar)')
addBars(thisone=8, bar0=90, bar1=99, labcol='gray90', 
        text1='Corynebacterium sundsvallense', 
        text2='NR_026375.1 (98% similar)')
addBars(thisone=12, bar0=100, bar1=109, labcol='gray90', 
        text1='Propionibacterium acnes', 
        text2='NR_074675.1 (96% similar)')

# addBars(thisone=9, bar0=110, bar1=119)
# addBars(thisone=10, bar0=120, bar1=129)
# addBars(thisone=8, bar0=130, bar1=139)
# addBars(thisone=12, bar0=140, bar1=149)


# addBars(thisone=1, bar0=30, bar1=39, Lines=FALSE)
# addBars(thisone=2, bar0=40, bar1=49)
# addBars(thisone=3, bar0=50, bar1=59)
# addBars(thisone=4, bar0=60, bar1=69)

dev.off()



```



