# Personal Microbial Cloud

## Sequence Data Cleanup

* load ecology packages
* read in a few functions that will be used later

```{r}
library(vegan)
library(labdsv)
library(ape) 
source('../data/pb_functions_for_manuscript.R')
```


Load data: 

* Classic OTU table. Manually remove hash character from first line. This also has taxonomy from GreenGenes database as last field. 
* Mapping file used in QIIME. Also removed hash from first line and removed comment line. 
* And then jive the order of samples in the OTU table and mapping file.


```{r}
pb <- read.delim('../data/otu_table_r.txt', head=TRUE, row.names=1)
pb.tax <- pb[, ncol(pb)]
pb <- pb[, -ncol(pb)]
pb <- t(pb) 
 
pb.map <- read.delim('../data/pb_map.txt', head=TRUE, row.names=1)

all(row.names(pb.map) %in% row.names(pb))
dim(pb.map); dim(pb)

pb.map <- pb.map[row.names(pb), ]
dim(pb); dim(pb.map); length(pb.tax)
identical(row.names(pb), row.names(pb.map))
```


Keep original copy before changing things. 

```{r}
pb.original <- pb 
pb.tax.original <- pb.tax
pb.map.original <- pb.map
```



```
# saved just in case of screwup. 
pb <- pb.original
pb.tax <- pb.tax.original
pb.map <- pb.map.original
```

Take out plant sequences

```{r}
plants <- grep('Streptophyt', pb.tax)
pb <- pb[, -plants]
pb.tax <- pb.tax[-plants]
rm(plants) 
```



Set up taxonomy table

```{r}
taxo <- makeTaxo(taxo.in=pb.tax, otu.table=pb)
dim(taxo); dim(pb) 
```

The big dataset started with `r sum(pb)` total sequences from `r ncol(pb)` OTUs and `r nrow(pb)` samples.


Identify those samples not used in the study. 

```{r}
nonexp <- which(pb.map$location == 'east' | 
  pb.map$location == 'west' | 
  pb.map$location == 'out')
pb <- pb[-nonexp, ]
pb.map <- pb.map[-nonexp, ]
rm(nonexp) 
dim(pb); dim(pb.map)
```



Index for controls, and clean up mapping table. 

```{r}
pb.map <- pb.map[, c('person', 'location', 'duration', 'sampleType')]
control <- which(pb.map$location == 'control')
pb <- pb[-c(control), ]
pb.map <- pb.map[-c(control), ]
dim(pb); dim(pb.map) 
identical(row.names(pb), row.names(pb.map))
```





Remove 4 contaminants that were introduced by PCR reagents. 
And remove control samples. 
This is detailed in `pb_analysis_Redo.Rmd`. 

```{r}
contaminants <- c('190873', '64285', '96443', '129992')
contaminants.index <- which(colnames(pb) %in% contaminants)
dim(pb); dim(pb.map)
pb <- pb[, -contaminants.index]
taxo <- taxo[-contaminants.index, ]  # pb
rm(control, contaminants, contaminants.index)
identical(row.names(pb.map), row.names(pb))
dim(pb.map); dim(pb); dim(taxo) 
```



Rarefy to 1000 seqs per sample. 

```{r}
sort(rowSums(pb))
pb <- pb[-which(rowSums(pb) < 1000), ]
pb.1000 <- rrarefy(pb, 1000)
pb.1000 <- pb.1000[, -which(colSums(pb.1000) == 0)]
pb.map <- pb.map[row.names(pb.1000), ]
taxo <- taxo[colnames(pb.1000), ]
dim(pb.1000); dim(pb.map); dim(taxo) 
```



Put them all in alphabetical order, since that gives nice groupings. 
Yay for labelling schemes!

```{r}
pb.map <- pb.map[order(row.names(pb.map)), ]
pb.1000 <- pb.1000[row.names(pb.map), ] 
```


Fix the abundance column - it still has abundance from full dataset, but abundance from rarefied dataset more meaningful. 

```{r}
identical(row.names(taxo), colnames(pb.1000))  # just to make sure
taxo$abundance <- colSums(pb.1000)
```


A few metrics to reference later.

```{r}
total.seqs <- sum(pb)
total.samples <- nrow(pb)
total.otus <- ncol(pb) 

rarefied.seqs <- sum(pb.1000)
rarefied.samples <- nrow(pb.1000)
rarefied.otus <- ncol(pb.1000)
```




Clean up factor levels. 

```{r}
pb.map$person <- factor(pb.map$person)
pb.map$location <- factor(pb.map$location)
pb.map$duration <- factor(pb.map$duration)
pb.map$sampleType <- factor(pb.map$sampleType)
pb.map$occ.person <- factor(paste(pb.map$location, pb.map$person, sep=''))  
 
pb.map$occ.person2 <- as.character(pb.map$occ.person)
pb.map$occ.person2[pb.map$location == "unocc"] <- "unocc"
pb.map$occ.person2 <- factor(pb.map$occ.person2)
```


How do samples fall into groups? I.e., do we have balance in treatments?

```{r}
table(pb.map$person, pb.map$sampleType)
table(pb.map$location)
table(pb.map$duration)
table(pb.map$sampleType) 
```


Indexing lists for quick parsing. 

```{r}
s22of <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
  pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter' 
	] 
s24of <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter'
	]
s22op <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s24op <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s22uf <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s24uf <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s22up <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]
s24up <- c(1:nrow(pb.map))[pb.map$person == 's2' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]

##
s32of <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter'
	]
s34of <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter'
	]
s32op <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s34op <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s32uf <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s34uf <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s32up <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]
s34up <- c(1:nrow(pb.map))[pb.map$person == 's3' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]

## 
s12of <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter'
	]
s14of <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'filter'
	]
s12op <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 2 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s14op <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 4 &
	pb.map$location == 'occ' &
	pb.map$sampleType == 'petri.dish'
	]
s12uf <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s14uf <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'filter'
	]
s12up <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 2 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]
s14up <- c(1:nrow(pb.map))[pb.map$person == 's1' & 
	pb.map$duration == 4 &
	pb.map$location == 'unocc' &
	pb.map$sampleType == 'petri.dish'
	]


## A few groupings to call up later

groups <- list(s22of, s32of, s12of, s22op, s32op, s12op, 
			   s22uf, s32uf, s12uf, s22up, s32up, s12up,
			   s24of, s34of, s14of, s24op, s34op, s14op,
			   s24uf, s34uf, s14uf, s24up, s34up, s14up)

names(groups) <- c('s22of', 's32of', 's12of', 's22op', 's32op', 's12op', 
			   's22uf', 's32uf', 's12uf', 's22up', 's32up', 's12up',
			   's24of', 's34of', 's14of', 's24op', 's34op', 's14op',
			   's24uf', 's34uf', 's14uf', 's24up', 's34up', 's14up')

gr2 <- list(s14of, s12of, s14op, s12op, s14uf, s12uf, s14up, s12up, 
			s34of, s32of, s34op, s32op, s34uf, s32uf, s34up, s32up, 
			s24of, s22of, s24op, s22op, s24uf, s22uf, s24up, s22up)

gr.name <- c('s14of', 's12of', 's14op', 's12op', 's14uf', 's12uf', 's14up', 's12up', 
			's34of', 's32of', 's34op', 's32op', 's34uf', 's32uf', 's34up', 's32up', 
			's24of', 's22of', 's24op', 's22op', 's24uf', 's22uf', 's24up', 's22up')

# clean up group names
rm(list=ls()[which(ls() %in% names(groups))])
```




Now save the image that can be used for analysis and figures. 

```{r}
rm(pb.original, pb.map.original, pb.tax.original, pb, pb.tax)
save.image('../data/pb_data.RData')
ls()  
```




