---
  output: html_document
---
  
  
```{r config}
library(knitr)
opts_chunk$set(background='gray80', tidy=FALSE, cache=FALSE)
```


cutoffs = 0.3, 0.5, 1, 2.5, 5, 10+ 


```{r readData}
calibrate <- read.delim('calibrate/combined.txt', sep='\t', head=TRUE)
part <- read.delim('subjectCounts/particlesCombined.txt', head=TRUE, sep='\t')
names(part)
```





# Calibrate particle loss ratio. 



```{r checkDateTime}
identical(calibrate$dataTimeSupply, calibrate$dataTimeInside)
dateTime <- calibrate$dataTimeSupply
ach <- calibrate$ACH
```


## Convert from .1ft3 to Liters per minute

```{r convertToLiters}
head(calibrate)
calibrate <- calibrate[, c(2:7, 9:14)]/2.83
calibrate$dateTime <- dateTime
calibrate$ach <- ach
```



```{r subsetCalibrate}
ach0 <- calibrate[calibrate$ach == 0, ]
ach1 <- calibrate[calibrate$ach == 1, ]
ach3 <- calibrate[calibrate$ach == 3, ]
dim(ach0); dim(ach1); dim(ach3)
```



## 0 ACH

The empty chamber gives the same 10+um particle counts from both counters. That means we can consider them representative and move on to calculate particle loss ratio in air change rate treatments. 

```{r plotCalCorrelations}
mean(ach0$Ch6DiffInside - ach0$Ch6DiffSupply)
boxplot(ach0$Ch6DiffInside - ach0$Ch6DiffSupply)
t.test(ach0$Ch6DiffInside - ach0$Ch6DiffSupply)
```


Somewhat different

```{r plot510CalCorrelations}
mean(ach0$Ch5DiffInside - ach0$Ch5DiffSupply)
boxplot(ach0$Ch5DiffInside - ach0$Ch5DiffSupply)
t.test(ach0$Ch5DiffInside - ach0$Ch5DiffSupply)
```

Way different. 

```{r plot35CalCorrelations}
mean(ach0$Ch4DiffInside - ach0$Ch4DiffSupply)
boxplot(ach0$Ch4DiffInside - ach0$Ch4DiffSupply)
t.test(ach0$Ch4DiffInside - ach0$Ch4DiffSupply)
```


Way way different. 

```{r plot13CalCorrelations}
mean(ach0$Ch3DiffInside - ach0$Ch3DiffSupply)
boxplot(ach0$Ch3DiffInside - ach0$Ch3DiffSupply)
t.test(ach0$Ch3DiffInside - ach0$Ch3DiffSupply)
```





## 1 ACH

Check for time lag correlation. 

3 minute is best but still weak. 

```{r corTimeLag1ACHch6}
cor.test(ach1$Ch6DiffInside, ach1$Ch6DiffSupply)
cor.test(ach1$Ch6DiffInside[-1], ach1$Ch6DiffSupply[-91])
cor.test(ach1$Ch6DiffInside[-c(1:2)], ach1$Ch6DiffSupply[-c(90:91)])
cor.test(ach1$Ch6DiffInside[-c(1:3)], ach1$Ch6DiffSupply[-c(89:91)])  # best fit but weak.
cor.test(ach1$Ch6DiffInside[-c(1:4)], ach1$Ch6DiffSupply[-c(88:91)])
```


Best fit is 4 minute lag. 

```{r corTimeLag1ACHch5}
cor.test(ach1$Ch5DiffInside, ach1$Ch5DiffSupply)
cor.test(ach1$Ch5DiffInside[-1], ach1$Ch5DiffSupply[-91])
cor.test(ach1$Ch5DiffInside[-c(1:2)], ach1$Ch5DiffSupply[-c(90:91)])
cor.test(ach1$Ch5DiffInside[-c(1:3)], ach1$Ch5DiffSupply[-c(89:91)])  
cor.test(ach1$Ch5DiffInside[-c(1:4)], ach1$Ch5DiffSupply[-c(88:91)])  # best fit
cor.test(ach1$Ch5DiffInside[-c(1:5)], ach1$Ch5DiffSupply[-c(87:91)])
```


No good fit. 

```{r corTimeLag1ACHch4}
cor.test(ach1$Ch4DiffInside, ach1$Ch4DiffSupply)
cor.test(ach1$Ch4DiffInside[-1], ach1$Ch4DiffSupply[-91])
cor.test(ach1$Ch4DiffInside[-c(1:2)], ach1$Ch4DiffSupply[-c(90:91)])
cor.test(ach1$Ch4DiffInside[-c(1:3)], ach1$Ch4DiffSupply[-c(89:91)])  
cor.test(ach1$Ch4DiffInside[-c(1:4)], ach1$Ch4DiffSupply[-c(88:91)])  
cor.test(ach1$Ch4DiffInside[-c(1:5)], ach1$Ch4DiffSupply[-c(87:91)])  
```


Decent fit with 4 minutes. 

```{r corTimeLag1ACHch3}
cor.test(ach1$Ch3DiffInside, ach1$Ch3DiffSupply)
cor.test(ach1$Ch3DiffInside[-1], ach1$Ch3DiffSupply[-91])
cor.test(ach1$Ch3DiffInside[-c(1:2)], ach1$Ch3DiffSupply[-c(90:91)])
cor.test(ach1$Ch3DiffInside[-c(1:3)], ach1$Ch3DiffSupply[-c(89:91)])  
cor.test(ach1$Ch3DiffInside[-c(1:4)], ach1$Ch3DiffSupply[-c(88:91)])  
cor.test(ach1$Ch3DiffInside[-c(1:5)], ach1$Ch3DiffSupply[-c(87:91)])  
```






## Find loss rate ratio. 

```{r}
zeros <- which(ach1$Ch6DiffSupply == 0 | ach1$Ch6DiffInside == 0)
ch6ach1LRR <- mean(ach1$Ch6DiffSupply[-zeros] / ach1$Ch6DiffInside[-zeros])
boxplot(ach1$Ch6DiffInside - ach1$Ch6DiffSupply)
boxplot(ach1$Ch6DiffInside - ach1$Ch6DiffSupply/ch6ach1LRR)
t.test(ach1$Ch6DiffInside - ach1$Ch6DiffSupply)
t.test(ach1$Ch6DiffInside - ach1$Ch6DiffSupply/ch6ach1LRR)
```




```{r}
# zeros <- which(ach1$Ch5DiffSupply == 0 | ach1$Ch5DiffInside == 0)
ch5ach1LRR <- mean(ach1$Ch5DiffSupply / ach1$Ch5DiffInside)
boxplot(ach1$Ch5DiffInside - ach1$Ch5DiffSupply/ch5ach1LRR)
boxplot(ach1$Ch5DiffInside - ach1$Ch5DiffSupply)
t.test(ach1$Ch5DiffInside - ach1$Ch5DiffSupply)
t.test(ach1$Ch5DiffInside - ach1$Ch5DiffSupply/ch5ach1LRR)
```

Check Channel 5 with 4 minute lag.  -  only a marginal difference. Maybe not worth considering. 

```{r}
mean(ach1$Ch5DiffInside[-c(1:4)] - ach1$Ch5DiffSupply[-c(88:91)])
sd(ach1$Ch5DiffInside[-c(1:4)] - ach1$Ch5DiffSupply[-c(88:91)])
boxplot(ach1$Ch5DiffInside[-c(1:4)] - ach1$Ch5DiffSupply[-c(88:91)])
t.test(ach1$Ch5DiffInside[-c(1:4)] - ach1$Ch5DiffSupply[-c(88:91)])
```





Channel 4. 

```{r}
ch4ach1LRR <- mean(ach1$Ch4DiffSupply / ach1$Ch4DiffInside)
boxplot(ach1$Ch4DiffInside - ach1$Ch4DiffSupply)
boxplot(ach1$Ch4DiffInside - ach1$Ch4DiffSupply/ch4ach1LRR)
t.test(ach1$Ch4DiffInside - ach1$Ch4DiffSupply)
t.test(ach1$Ch4DiffInside - ach1$Ch4DiffSupply/ch4ach1LRR)
```

Channel 4 with time lag. No difference in variance - tried several times. 

```{r}
mean(ach1$Ch4DiffInside[-c(1:5)] - ach1$Ch4DiffSupply[-c(87:91)])
sd(ach1$Ch4DiffInside[-c(1:5)] - ach1$Ch4DiffSupply[-c(87:91)])
boxplot(ach1$Ch4DiffInside - ach1$Ch4DiffSupply)
t.test(ach1$Ch4DiffInside - ach1$Ch4DiffSupply)
```



Channel 3. 

```{r}
ch3ach1LRR <- mean(ach1$Ch3DiffSupply / ach1$Ch3DiffInside)
boxplot(ach1$Ch3DiffInside - ach1$Ch3DiffSupply)
boxplot(ach1$Ch3DiffInside - ach1$Ch3DiffSupply/ch3ach1LRR)
t.test(ach1$Ch3DiffInside - ach1$Ch3DiffSupply)
t.test(ach1$Ch3DiffInside - ach1$Ch3DiffSupply/ch3ach1LRR)
```






















# Real particle counts



```{r}
part[, c(4:15, 17:28)] <- part[, c(4:15, 17:28)]/2.8
part$subTreat <- factor(paste0(part$subject, part$treatment))
part$diff10 <- part$Ch6DiffInside - part$Ch6DiffSupply/ch6ach1LRR
part$diff510 <- part$Ch5DiffInside - part$Ch5DiffSupply/ch5ach1LRR
part$diff35 <- part$Ch4DiffInside - part$Ch4DiffSupply/ch4ach1LRR
part$diff13 <- part$Ch3DiffInside - part$Ch3DiffSupply/ch3ach1LRR
```



3 minute delay?

```{r, eval=TRUE}
plot(jitter(c(0, part$Ch6DiffInside)) ~ jitter(c(part$Ch6DiffSupply, 0)))
segments(0,0,100,100)
cor.test(c(0, part$Ch6DiffInside), c(part$Ch6DiffSupply, 0))


test <- part[part$subTreat == 's8low', ]

cor.test(test$Ch6DiffInside, test$Ch6DiffSupply)
plot(jitter(test$Ch6DiffInside), jitter(test$Ch6DiffSupply))

cor.test(test$Ch6DiffInside[2:90], test$Ch6DiffSupply[1:89])
plot(jitter(test$Ch6DiffInside[2:90]), jitter(test$Ch6DiffSupply[1:89]))

cor.test(test$Ch6DiffInside[3:90], test$Ch6DiffSupply[1:88])
plot(jitter(test$Ch6DiffInside[3:90]), jitter(test$Ch6DiffSupply[1:88]))

cor.test(test$Ch6DiffInside[4:90], test$Ch6DiffSupply[1:87])
plot(jitter(test$Ch6DiffInside[4:90]), jitter(test$Ch6DiffSupply[1:87]))

cor.test(test$Ch6DiffInside[5:90], test$Ch6DiffSupply[1:86])
plot(jitter(test$Ch6DiffInside[5:90]), jitter(test$Ch6DiffSupply[1:86]))

##
plot(jitter(c(0, part$Ch5DiffInside)) ~ jitter(c(part$Ch5DiffSupply, 0)))
segments(0,0,100,100)
cor.test(c(0, part$Ch5DiffInside), c(part$Ch5DiffSupply, 0))

cor.test(test$Ch5DiffInside, test$Ch5DiffSupply)
plot(jitter(test$Ch5DiffInside), jitter(test$Ch5DiffSupply))

cor.test(test$Ch5DiffInside[2:90], test$Ch5DiffSupply[1:89])
plot(jitter(test$Ch5DiffInside[2:90]), jitter(test$Ch5DiffSupply[1:89]))

cor.test(test$Ch5DiffInside[3:90], test$Ch5DiffSupply[1:88])
plot(jitter(test$Ch5DiffInside[3:90]), jitter(test$Ch5DiffSupply[1:88]))

cor.test(test$Ch5DiffInside[4:90], test$Ch5DiffSupply[1:87])
plot(jitter(test$Ch5DiffInside[4:90]), jitter(test$Ch5DiffSupply[1:87]))

cor.test(test$Ch5DiffInside[5:90], test$Ch5DiffSupply[1:86])
plot(jitter(test$Ch5DiffInside[5:90]), jitter(test$Ch5DiffSupply[1:86]))

```



```{r, eval=TRUE}
part <- part[, c('subTreat', 'diff10', 'diff510', 'diff35', 'diff13')]
```




```{r}
dim(part)
```




```{r}
cols <- rep('tomato', 10)
cols[c(1, 5, 6, 8)] <- 'cornflowerblue'
#males <- c(1, 5, 6, 8)
#females <- c(2, 3, 4, 7)
```



```{r}
for (i in 1:nlevels(part$subTreat)) {
	lev <- levels(part$subTreat)[i]
	assign(lev, part[part$subTreat == lev, ])
	}
	
list.Low <- ls(pattern = 'low')
list.High <- ls(pattern = 'high')

```




```{r}
for (i in 1:length(list.Low)) {
	print(dim(get(list.Low[i])))
	}
for (i in 1:length(list.High)) {
	print(dim(get(list.High[i])))
	}
```


```{r}
plot(s1low$diff10, type='l')
plot(s1high$diff10, type='l')
```


```{r}
## Low
Agg18 <- rep(letters[1:18], each = 5)
for (i in 1:length(list.Low)) {
	name <- paste("agg.", list.Low[i], sep = "")
	assign(name, aggregate(get(list.Low[i])[, -1], by = list(Agg18), FUN = "sum"))
}

Agg.df <- data.frame(matrix(0, 0, 5))
Aggs <- ls(pattern = "agg.")
for (i in 1:length(Aggs)) {
	Agg.tmp <- get(Aggs[i])
	# print(get(Aggs[i]))
	Agg.df <- data.frame(rbind(Agg.df, 
		data.frame(rep(letters[i], nrow(Agg.tmp)), 
			Agg.tmp[, 2:5])))

}
names(Agg.df)[1] <- 'group'
# Agg.df

rm(list=ls(pattern='agg.'))
```


```{r}


## High
Agg18 <- rep(letters[1:18], each = 5)
for (i in 1:length(list.High)) {
	name <- paste("agg.", list.High[i], sep = "")
	assign(name, aggregate(get(list.High[i])[, -1], by = list(Agg18), FUN = "sum"))
}

Agg.df.High <- data.frame(matrix(0, 0, 5))
Aggs <- ls(pattern = "agg.")
for (i in 1:length(Aggs)) {
	Agg.tmp <- get(Aggs[i])
	# print(get(Aggs[i]))
	Agg.df.High <- data.frame(rbind(Agg.df.High, 
		data.frame(rep(letters[i], nrow(Agg.tmp)), 
			Agg.tmp[, 2:5])))

}
names(Agg.df.High)[1] <- 'group'
# Agg.df.High
```


```{r}
boxplot(Agg.df$diff10 ~ Agg.df$group, cex=.6, col=cols, pch=16, 
	las=1, xaxt='n', 
	ylab='')
mtext(expression(paste('Particles (>10', mu, 'm) L' ^{-1},  ' min' ^{-1})), 
	side=2, line=2)
abline(h=0, col='gray', lty=3, lwd=2)
mtext('People', side=1)
mtext(c('Men', 'Women'), side=1, at=c(3.5, 5.5), line=1.5, font=2, cex=1.4,
	col=c('cornflowerblue', 'tomato'))
```



```{r}
boxplot(Agg.df.High$diff10 ~ Agg.df.High$group, cex=.6, col=cols, pch=16, 
	las=1, xaxt='n', 
	ylab='')
mtext(expression(paste('Particles (>10', mu, 'm) L' ^{-1},  ' min' ^{-1})), 
	side=2, line=2)
abline(h=0, col='gray', lty=3, lwd=2)
mtext('People', side=1)
mtext(c('Men', 'Women'), side=1, at=c(3.5, 5.5), line=1.5, font=2, cex=1.4,
	col=c('cornflowerblue', 'tomato'))

```


```{r}


boxplot(Agg.df$diff510 ~ Agg.df$group, cex=.6, col=cols, pch=16, 
	las=1, xaxt='n', 
	ylab='')
mtext(expression(paste('Particles (>10', mu, 'm) L' ^{-1},  ' min' ^{-1})), 
	side=2, line=2)
abline(h=0, col='gray', lty=3, lwd=2)
mtext('People', side=1)
mtext(c('Men', 'Women'), side=1, at=c(3.5, 5.5), line=1.5, font=2, cex=1.4,
	col=c('cornflowerblue', 'tomato'))
```



```{r}
boxplot(Agg.df.High$diff510 ~ Agg.df.High$group, cex=.6, col=cols, pch=16, 
	las=1, xaxt='n', 
	ylab='')
mtext(expression(paste('Particles (>10', mu, 'm) L' ^{-1},  ' min' ^{-1})), 
	side=2, line=2)
abline(h=0, col='gray', lty=3, lwd=2)
mtext('People', side=1)
mtext(c('Men', 'Women'), side=1, at=c(3.5, 5.5), line=1.5, font=2, cex=1.4,
	col=c('cornflowerblue', 'tomato'))

```





```{r}
save(Agg.df, file='particleCounts.RData')
```

